On the Synchronization of Computational RFIDs

### 2018-05-14 response/response_letter.tex

\bigskip\bigskip

%\noindent{Marwan Krunz, EIC, and Kevin Almeroth, Associate EIC \\
%	IEEE Transactions on Mobile Computing
%}

\bigskip\bigskip

\begin{flushright}
Dr. Przemys{\l}aw Pawe{\l}czak \\
Embedded Software Group \\ 
Delft University of Technology \\ 
Van Mourik Broekmanweg 6\\
2628\,XE Delft, The Netherlands \\
E-mail: p.pawelczak@tudelft.nl\\
Phone: +31 15 278 7491\\
\end{flushright}

@@

\textbf{TMC-2017-11-0695:} {\sl On the Synchronization of Computational RFIDs}

\bigskip

Dear Editor,\\

We would like to thank for handling our submission and providing a rapid review results. We are also grateful to the reviewers for their substantial effort spent for evaluating our work. \\

With this letter we provide our response to all comments of the reviewers and the revised version of our paper. We look forward to hearing from you at the earliest and we would be happy to provide further clarifications and revisions if necessary. \\

### 2018-05-14 response/response_letter.tex

in the name of all Authors---Dr. Przemys{\l}aw Pawe{\l}czak

\pagebreak

\begin{editor}
\referee{The authors need to revise the paper accordingly and prepare for the second review. The reviewers mainly concerns in the following aspects.
	\begin{enumerate}
		\item Clearly explain the motivation, describe clearly about the problem definition, and describe what new challenges address in this paper;
		\item Address the technical details mentioned by Reviewers 1 and 2;
		\item More results to present mentioned by Reviewer 3.
\end{enumerate}
}

\response{Thank you once more for coordinating the review process of our paper. We hope that we have revised our article in accordance with the responses of the reviewers. We believe we have addressed all three points that the editor raised as accurately as possible. Briefly, in the revised article:
\begin{itemize}
	\item We provided better a motivation in Section 1.1, we mentioned our contributions more clearly by also citing the related work suggested by Reviewer 1 in Section 2.2.2;
	\item We addressed all of the technical details, especially the differences of our work from slot and bit-level synchronization in EPC Gen2 standard in Section 2.2.2;
	\item We clarified the misunderstanding about multi-hop CRFID networks by clearly stating in Section 1.1. that these networks do not exist yet and the focus of our work is current of-the-shelf CRFID and reader platforms.
\end{itemize}

We request the editor to refer to our respective responses below; in particular \hyperref[R1:1]{[R1:1]}, \hyperref[R1:2]{[R1:2]}, \hyperref[R1:3]{[R1:3]}, \hyperref[R2:1]{[R2:1]} and \hyperref[R3:2]{[R3:2]}. Additions with respect to the previous version of our submission are marked in blue text in our latest paper version.}

### 2018-05-14 response/response_letter.tex

\referee{Also, there are another view to classify the synchronization problems: slot level synchronization (time frame is divided into slots) or bit level synchronization. Slot level synchronization has been studied for decades and is well addressed. The bit-level synchronization is not well investigated yet and attracts much attention from academic communities, e.g., the following three references. The authors should discuss and clarify whether they study the bit-level synchronization mentioned in these references.}

\response{We thank the reviewer for this comments and for pointing us to these relevant papers, which we cite in Section 2.2.2 of our revised work. Thanks to the comment of the reviewer we have noticed that in the previous version of our submission we did not highlighted the main difference between our work and the works pertaining to the \emph{bit-level synchronization}.

To clarify this difference: as indicated earlier by the reviewer, it is possible to exploit modulation properties of signals from the reader to the tags (such as bit timing in FM0 modulation) in order to get the timing measurements of the regular patterns in the signal, i.e. bit-level synchronization. However, the objective of our work has the following fundamental differences as compared to these studies:

\begin{itemize}
	\item \textbf{Portability:} Bit-level synchronization relies on high-precision clock to obtain microsecond-precision timing measurements. Despite, in our work, we used the external 32\,kHz clock that can run in low-power standby mode and the \emph{higher-level} commands of the EPC Gen2 standard, rather than exploiting its low-level signal properties. Therefore, our approach is independent of how EPC Gen2 is implemented---providing a level of portability across different platforms. However, our synchronization algorithms can also be integrated into the bit-level synchronization;

@@

	\item \textbf{Explicit Clock:} Bit-level synchronization does not output an explicit clock value and it does not compensate for the clock drift. Rather, it uses offset compensation only and \emph{guard time} to reduce the effect of the clock drift. On the contrary, we provided time synchronization algorithms that compensate for the clock drift and provide an \emph{explicit} common time notion to CRFID applications;

	\item \textbf{Energy Budget:} Due to marginal energy budgets, we proposed synchronization algorithms that require minimum number of computation steps and memory space;

	\item \textbf{Power Failures:} Providing synchronization service to the application layer requires to cope with power failures, which has not been investigated so far. To address this issue, we provided synchronization algorithms that can recover easily after power failures.
\end{itemize}

We would like to emphasize that none of these issues have been addressed and studied so far within the context of CRFIDs. Again, the above discussion has been briefly added to Section 2.2.2.}

@@

\referee{What if WISP tags are simultaneously covered by multiple readers. In this case, the readers are not synchronized themselves and may have different time-stamps. Then, how does the proposed method work?}

\response{As we indicated in the introduction of the article (Section 1) there are already existing synchronization techniques that provide synchronization among RFID readers. Therefore, if there are multiple readers that cover multiple WISP tags, first the readers should be synchronized with each other and then the tags will be synchronized to the RFID readers. This approach will cause any danger to the protocol we designed. We mentioned this explicitly in Section 1.1 and Section 7.5 of our revised paper.}

@@

\referee{The novelty of this paper is relatively limited because the proposed event-based reader-tag synchronization method is originated from previous literature. Moreover, the main part of the theoretical analytics is also from previous work [24,31]. Although the authors claimed that they use a different clock drift model in this article, i.e. Brownian motion rather than uniform distribution, I still this significantly reduces the theoretical depth of this paper. The authors should show what NEW challenges beyond previous work are addressed by this paper.}

\response{As we indicated in our previous responses (see \hyperref[R1:1]{[R1:1]} and \hyperref[R1:2]{[R1:2]}), our article has the following novelties over the previous results:
\begin{enumerate}
	\item \textbf{Difference with bit-level synchronization:} Bit-level synchronization does not compensate for clock drift and does not provide explicit time notion to the tags. In order to output a synchronized clock value, a time synchronization algorithm is required;
	\item \textbf{Energy budget of synchronization algorithm:} Since the energy budget of the tags are limited, synchronization approaches require marginal number of computation and memory. This is challenging, which was our main motivation for applying a PI-controller based synchronization for the first time in this domain;
	\item \textbf{Consideration of intermittent power supply:} Maintaining an explicit synchronized clock despite power failures has never been explored before in the field of CRFIDs. We also provided how our synchronization technique overcomes frequent power failures and provide a stable synchronized clock value;
	\item \textbf{Novel theoretical consideration:} We also provided an in-depth theoretical analysis of the proposed synchronization techniques. As indicate by the reviewer, the main part of the analysis indeed originates from the previous work. However, the previous analysis that assumes uniform distribution for the clock drifts was not realistic---the reason we changed the analysis to use the Brownian motion. This modification did not change the proof structure but it changed the individual steps of the proof---that led to \emph{different steady-state error expressions}.
\end{enumerate}}

### 2018-05-14 response/response_letter.tex

\end{responses}

\begin{responses}

\referee{It is better to simplify the descriptions of the challenges and existing works, especially some contents that are not directly related to the synchronization.}

\response{To address this comment of the reviewer, we modified the description in Section 2. That is, we have listed novelties of our contribution in the context of new related work in Section 2.2.2. Accompanying this revision, in Section 1 we have provided a better motivation.}

@@

\referee{It is better to report some results or discuss the performance of the proposed approach in the multi-hop networks, since the authors highlights the advantages of multi-hop networks several times in the introduction.}

\response{We thank the reviewer for this comment. However, we feel that there has been a misunderstanding regarding the concept of backscatter (CRFID) multi-hop networks. As of now, CRFID tags cannot operate in a multi-hop manner since they can only communicate directly with the RFID readers. In the introduction of our article (Section 1.1), we only stated that we \emph{anticipate} that after enabling tag-to-tag backscatter communication, multi-hop CRFID networks will \emph{emerge} and therefore time synchronization will be more important. However, as of now, we do not have access to (nor we are not aware of) any network infrastructure to demonstrate neither multi-hop communication nor multi-hop synchronization.

Our work applies to the case when several RFID readers cover large tag population simultaneously. Then the whole network can be synchronized by first allowing RFID readers to access central reference time, e.g. using the Internet; then individual tags can be synchronized to the individual RFID readers. To clarify the connection of our work to the multi-hop networks, we have provided an extra clarifying sentences in Section 1.1.}

### 2018-05-15 response/response_letter.tex

in the name of all Authors---Dr. Przemys{\l}aw Pawe{\l}czak

\pagebreak

\begin{editor}
\referee{The authors need to revise the paper accordingly and prepare for the second review. The reviewers mainly concerns in the following aspects.
	\begin{enumerate}
		\item Clearly explain the motivation, describe clearly about the problem definition, and describe what new challenges address in this paper;
		\item Address the technical details mentioned by Reviewers 1 and 2;
		\item More results to present mentioned by Reviewer 3.
\end{enumerate}
}

\response{Thank you once more for coordinating the review process of our paper. We hope that we have revised our article in accordance with the responses of the reviewers. We believe we have addressed all three points that the editor raised as accurately as possible. Briefly, in the revised article:
\begin{itemize}
	\item We provided a better motivation in Section 1.1, we mentioned our contributions more clearly by also citing the related work suggested by Reviewer 1 in Section 2.2.2;
	\item We addressed all of the technical details, especially the differences of our work from slot and bit-level synchronization in EPC Gen2 standard in Section 2.2.2;
	\item We clarified the misunderstanding about multi-hop CRFID networks by clearly stating in Section 1.1 that these networks do not exist yet and the focus of our work is on the current of-the-shelf CRFID and reader platforms.
\end{itemize}

### 2018-05-15 response/response_letter.tex

\begin{itemize}
	\item \textbf{Portability:} Bit-level synchronization relies on high-precision clock to obtain microsecond-precision timing measurements. Despite, in our work, we used the external 32\,kHz clock that can run in low-power standby mode and the \emph{higher-level} commands of the EPC Gen2 standard, rather than exploiting its low-level signal properties. Therefore, our approach is independent of how EPC Gen2 is implemented---providing a level of portability across different platforms. However, our synchronization algorithms can also be integrated into the bit-level synchronization;

	\item \textbf{Explicit Clock:} Bit-level synchronization does not output an explicit clock value and it does not compensate for the clock drift. Rather, it uses offset compensation only and \emph{guard time} to reduce the effect of the clock drift. On the contrary, we provided time synchronization algorithms that compensate for the clock drift and provide an \emph{explicit} common time notion to CRFID applications;

	\item \textbf{Energy Budget:} Due to marginal energy budgets, we proposed synchronization algorithms that require minimum number of computation steps and memory space;

	\item \textbf{Power Failures:} Providing synchronization service to the application layer requires coping with power failures, which has not been investigated so far. To address this issue, we provided synchronization algorithms that can recover easily after power failures.
\end{itemize}

@@

We would like to emphasize that none of these issues have been addressed and studied so far within the context of CRFIDs. Again, the above discussion has been briefly added to Section 2.2.2.}

\referee{What if WISP tags are simultaneously covered by multiple readers. In this case, the readers are not synchronized themselves and may have different time-stamps. Then, how does the proposed method work?}

\response{As we indicated in the introduction of the article (Section 1) there are already existing synchronization techniques that provide synchronization among RFID readers. Therefore, if there are multiple readers that cover multiple WISP tags, first the readers should be synchronized with each other and then the tags will be synchronized to the RFID readers. This approach will not cause any danger to the protocol we designed. We mentioned this explicitly in Section 1.1 and Section 7.5 of our revised paper.}

### 2018-05-15 response/response_letter.tex

\referee{Also, there are another view to classify the synchronization problems: slot level synchronization (time frame is divided into slots) or bit level synchronization. Slot level synchronization has been studied for decades and is well addressed. The bit-level synchronization is not well investigated yet and attracts much attention from academic communities, e.g., the following three references. The authors should discuss and clarify whether they study the bit-level synchronization mentioned in these references.}

\response{We thank the reviewer for this comments and for pointing us to these relevant papers, which we cite in Section 2.2.2 of our revised work. Thanks to the comment of the reviewer we have noticed that in the previous version of our submission we did not highlighted the main difference between our work and the works pertaining to the \emph{bit-level synchronization}.

To clarify this difference: as indicated earlier by the reviewer, it is possible to exploit modulation properties of signals from the reader to the tags (such as bit timing in FM0 modulation) in order to get the timing measurements of the regular patterns in the signal, i.e. bit-level synchronization. However, the objective of our work has the following fundamental differences as compared to these studies:

\begin{itemize}
	\item \textbf{Portability:} Bit-level synchronization relies on a high-precision clock to obtain microsecond-precision timing measurements. However, rather than exploiting low-level signal properties, in our work we rely on (i) external 32\,kHz clock that can run in low-power standby mode and (ii) the \emph{higher-level} commands of the EPC Gen2 standard. Therefore, our approach is independent of how EPC Gen2 is implemented---providing a level of portability across different platforms. We emphasize that our synchronization algorithms can also be integrated into the bit-level synchronization;

### 2018-05-17 TMC-2017-11-0695-main.tex
figure
Figure 1 size reduced
typo
latest changes

\usepackage{algorithm,setspace}
\usepackage[noend]{algpseudocode} % needed for: algorithm pseudocode
\usepackage{color}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{soul}
\usepackage{cancel}
\usepackage{listings}
\usepackage{tikz} % needed for: tikz pics
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.pathreplacing}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\columnwidth]{figures/scenario.pdf}
	\caption{\label{fig:motivation}\color{blue}A representative greenhouse monitoring application using CRFIDs. Two readers are deployed to provide power to the individual tags and to collect their periodic sensor readings.}
\end{figure}

@@

The focus of this article is to answer the question of \emph{how to design a building block that synchronizes intermittently-powered CRFID systems?} To this end, we investigate the WISP platform and provide initial observations and limitations pertaining to the synchronization of these devices. In particular, the main contributions and findings of 
this article are:
%
\begin{itemize}
	\item To the best of our knowledge, this is the first study that focuses on the synchronization of tags in CRFID systems. Even though there are studies on the synchronization of multiple RFID readers, e.g.~\cite{reader_sync}, we are unaware of any study that provides synchronization of individual tags in the communication range of an RFID reader. 
	\item By addressing the aforementioned challenges, we design and implement \emph{the first synchronization primitive} for the tags, namely \emph{WISP-Sync}, inspired by the Proportional-Integral (PI) controllers~\cite{pi2015,yildirim2017adaptive}. WISP-Sync (i) 
	requires only a few computation steps to run efficiently under limited harvested energy and keeps a few variables to hold the synchronization state to recover from power interruptions with minimum overhead---\emph{addressing Challenge I}; (ii) is adaptive to react to clock instabilities in a fast manner---\emph{addressing Challenge II}.
	\item \add{Our time synchronization algorithms compensate for the clock drift and enable the first step to enable an \emph{explicit clock} for CRFID applications. Moreover, our implementations are independent of how EPC Gen2 is implemented---providing a level of portability across different platforms.}
	%
	\item We provide theoretical analysis to prove that this primitive establishes synchronization and to reveal the factors affecting its synchronization performance.
	%
	\item Evaluations in our testbed show that an average synchronization error of approximately 1\,ms can be ensured among tags with an energy overhead of 1.85\,$\mu$J per synchronization round using this  primitive.
\end{itemize}
%
We note that WISP-side implementations throughout the article are done using C and assembly language. All source codes and scripts used to generate the results in the article (including parsing, post-processing and measurement results) are available via~\href{https://goo.gl/RuWIZ2}{https://goo.gl/RuWIZ2}.

@@

Let $t_0$ denote the first event reception time at the tag. Using the pre-measured event period $\tau$ of the RFID reader, the tag can estimate the next event reception time at time $t_1$ as
\begin{eqnarray}
\hat{C}_w(t_1) &=& C_w(t_0)+\tau(1+\hat{f}_w^r(t_0)), \label{eq:core}
\end{eqnarray}
where $\hat{f}_w^r(t_0)$ denotes the \emph{estimated relative frequency} at time $t_0$, which is an estimate for \eqref{eq:relative_frequency}. The intuition behind \eqref{eq:core} is that $1+\hat{f}_w^r(t_0)$ gives the estimated value of $f_r(t_0)/f_w(t_0)$, $\tau f_r(t_0)$ gives the estimated local time passed on the reader side and $\tau f_r(t_0)/f_w(t_0)$ gives the estimated number of clock ticks on the tag side that will pass until the next packet reception. Redefining \eqref{eq:gamma_error}, the \emph{prediction error} at time $t_1$ can be written as
\begin{eqnarray}
\gamma(t_1) & \!\!\!\!\!=\!\!\!\!\!\! & 
C_{w}(t_{1}\!)\!-\!\hat{C}_{w}(t_{1}\!) \!=\! 
C_{w}(t_{1}\!)\!-\!C_{w}(t_{0}\!)\!-\!\tau(1\!\!+\!\!\hat{f}_{w}^{r}(t_0\!)). \label{eq:gamma_t_1}
\end{eqnarray}
Apparently, the objective is to minimize the prediction error so that the event reception time is estimated correctly. As can be observed from \eqref{eq:gamma_t_1}, the estimation error on the relative clock frequency at time $t_0$, i.e. $\hat{f}_w^r(t_0)$, depends on the mean event period $\tau$, and its variance (since they affect $C_w(t_1)-C_w(t_0)$). Inspired by the integral controller update rule in \cite[Section III, (5)]{pi2015}, we propose to update the estimated relative frequency at time $t_1$ as
\begin{eqnarray}
\hat{f}_{w}^{r}(t_1) &=& \hat{f}_{w}^{r}(t_0)+\beta\gamma(t_1), 
\label{eq:update_rule}
\end{eqnarray}
where $\beta$ is the \emph{integral gain}. Assume that the uncertainty of the event period is $\varepsilon\sim\mathcal{N}(0,\sigma^2_{\tau})$ and the instantaneous change on the relative clock frequency is $\frac{df_{w}^{r}(t)}{dt}\sim\mathcal{N}(0,\sigma_{\eta}^{2})$. Under these assumptions, it can be proven that if $0<\beta<\frac{2}{\tau}$ is satisfied the synchronization will hold eventually; i.e. $\hat{f}_{w}^{r}=f_{w}^{r}$. Moreover, the mean and the variance ($\sigma^2_\gamma$) of the steady-state (asymptotic) estimation error of WISP-Sync can be given by 
\begin{eqnarray}
\gamma & \!\!\!\!\!\sim\!\!\!\!\! & \mathcal{N} \!\!\left(\!\!0,\sigma^2_\gamma\!\!=\!\!\frac{\sigma^2_{\eta}\!\left(\!\tau^2\!-\!\beta\tau^3\!+\!\frac{\beta^2\tau^4}{3}\!\right)\!\!+\!\!\beta^2\sigma^2_{\tau}\tau}{(2\beta-\beta^2\tau)}\!\!+\!\!\frac{\sigma^2_{\eta}\tau^3}{3}\!+\!\sigma^2_{\tau} \!\!\right),
\label{eq:sync_performance}
\end{eqnarray}
which is derived in Section~\ref{sec:app_b} of the Appendix.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/sim-theory.eps}
	\caption{\label{fig:theory-vs-simulations} The plot of the cumulative distribution function (CDF) of $\gamma$ in \eqref{eq:sync_performance} and the CDF of the steady-state $\gamma$ obtained during numerical simulations of \eqref{eq:update_rule} (by implementing \eqref{eq:theta} and \eqref{eq:delta_recursive} in Appendix). We used $\tau=1$, $\beta= 0.0001$, $\sigma_{\tau}=0.00005$ and our plots depict the theoretical and simulation results with respect to $\sigma_\eta = 0.0002$ and $\sigma_\eta = 0.0008.$ It can be observed that \eqref{eq:sync_performance} is consistent with the simulations and bigger uncertainty leads to bigger synchronization error.}
\end{figure}
In Fig.~\ref{fig:theory-vs-simulations}, we emphasize that our theoretical derivation in \eqref{eq:sync_performance} is consistent with the simulations of \eqref{eq:update_rule}. We refer again the reader to the Appendix for the detailed proofs of the convergence and the asymptotic estimation error. Moreover, in Section~\ref{sec:adaptibility}, we discuss the impact of the parameters $\tau$ and $\beta$ on the synchronization accuracy.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.4\columnwidth]{figures/alpha-period-vs-variance.eps}
	\caption{\label{fig:simulations} Numerical results showing the relations between event period, integral gain and the steady-state variance. First, we set $\sigma_{\tau}=0.4097$, $\sigma_\eta = 0.0008$, $\beta=1$ and plotted $\sigma^2_{\gamma}$ using different $\tau$ (left hand side). Observe that increasing $\tau$ increases $\sigma^2_{\gamma}$. Then, we set $\sigma_{\tau}=0.4097$,	$\tau=1$ and $\beta \in (0,2)$ due to \eqref{eq:step_size_bounds-pi}. Using different values for $\sigma_\eta$, we plotted how $\sigma^2_{\gamma}$ changes with respect to the $\beta$ (right hand side). Observe that for each value of $\sigma_\eta$, there is a particular $\beta$ that minimizes $\sigma^2_{\gamma}$.}
\end{figure*}

@@

Since we measured the event period as 226.76\,ms on average (see Fig.~\ref{fig:block_write_delay}), we set $\tau=7086$ 
clock ticks since the local clock of the tag is operating at 32\,kHz and each clock tick occurs every 32 microseconds. First, we explored how $\beta$ affects the performance of the algorithm. By substituting $\tau$ into \eqref{eq:step_size_bounds-pi}, the convergence condition becomes $0<\beta<0.000284$ in our case. We present the synchronization error with different $\beta$ values in Fig.~\ref{fig:beta}. Observe that the synchronization error tends to get smaller with smaller $\beta$ values but the convergence time increases negligibly. Moreover, due to the low-precision 32\,kHz clock, decreasing $\beta$ has no significant effect after some point. We mention that  in order to make WISP-Sync adapt faster, WISP-Sync can start with a big value of $\beta$ to converge fast and then decrease $\beta$ gradually to decrease the synchronization error, i.e. the adaptive strategy proposed in~\cite{pi2015}. However, from  Fig.~\ref{fig:beta}, we consider a fixed value of $\beta=0.0001$ for the next evaluation steps.

@@

Even though the event-based approach is relatively lightweight and adaptive as compared to the sender-receiver synchronization, it has some limitations. First, the tag is unable to obtain an \emph{explicit reference clock value}. The only variable it tunes is the $\hat{f}_w^r$ that represents the relative frequency of the software clock with respect to the clock of the reader. Therefore, we require additional steps that will allow tags to obtain the actual reference clock value. \add{Moreover, if the tags are covered by different readers, the first step should be to first synchronize the readers to the real-time. e.g. by allowing them to access time servers using internet.}
%However, to the best of our knowledge, there is no such command in the EPC Gen 2 standard that allows to send the explicit RFID reader time to the tag.
Second, WISP-Sync requires knowledge about $\tau$, which can be RFID reader dependent and requires a sniffer setup described in Section~\ref{sec:sniffer} to measure. This limitation can be addressed by employing an iterative estimation as an initial step to the WISP-Sync: the tag can start with an initial estimated $\tau$ value, and improve its estimation as it receives successive packets from the reader. After some iterations, the tag can have a rough estimate $\hat{\tau}$ that can be used on behalf of $\tau$. However, estimation errors might also perturb the synchronization accuracy. Last, the steps in Fig.~\ref{fig:event_based} present the communication scenario between one WISP and the reader. To allow other tags sniff this communication and synchronize, a broadcast primitive is crucial---in the next section we show how to overcome this limitation.

@@

\input{appendix_tmc}

>>>>

\usepackage{algorithm,setspace}
\usepackage[noend]{algpseudocode} % needed for: algorithm pseudocode
\usepackage{color}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
%\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{soul}
\usepackage{cancel}
\usepackage{listings}
\usepackage{tikz} % needed for: tikz pics
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.pathreplacing}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\columnwidth]{figures/scenario.pdf}
	\caption{\label{fig:motivation}\color{blue}A representative greenhouse monitoring application using CRFIDs. Two readers are deployed to provide power to the individual tags and to collect their periodic sensor readings.}
\end{figure}

@@

The focus of this article is to answer the question of \emph{how to design a building block that synchronizes intermittently-powered CRFID systems?} To this end, we investigate the WISP platform and provide initial observations and limitations pertaining to the synchronization of these devices. In particular, the main contributions and findings of 
this article are:
%
\begin{itemize}
	\item To the best of our knowledge, this is the first study that focuses on the synchronization of tags in CRFID systems. Even though there are studies on the synchronization of multiple RFID readers, e.g.~\cite{reader_sync}, we are unaware of any study that provides synchronization of individual tags in the communication range of an RFID reader. 
	\item By addressing the aforementioned challenges, we design and implement \emph{the first synchronization primitive} for the tags, namely \emph{WISP-Sync}, inspired by the Proportional-Integral (PI) controllers~\cite{pi2015,yildirim2017adaptive}. WISP-Sync (i) 
	requires only a few computation steps to run efficiently under limited harvested energy and keeps a few variables to hold the synchronization state to recover from power interruptions with minimum overhead---\emph{addressing Challenge I}; (ii) is adaptive to react to clock instabilities in a fast manner---\emph{addressing Challenge II}.
	\item \add{Our time synchronization algorithms compensate for the clock drift and enable the first step to enable an \emph{explicit clock} for CRFID applications. Moreover, our implementations are independent of how EPC Gen2 is implemented---providing a level of portability across different platforms.}
	%
	\item We provide theoretical analysis to prove that this primitive establishes synchronization and to reveal the factors affecting its synchronization performance.
	%
	\item Evaluations in our testbed show that an average synchronization error of approximately 1\,ms can be ensured among tags with an energy overhead of 1.85\,$\mu$J per synchronization round using this  primitive.
\end{itemize}
%
We note that WISP-side implementations throughout the article are done using C and assembly language. All source codes and scripts used to generate the results in the article (including parsing, post-processing and measurement results) are available via~\url{https://goo.gl/RuWIZ2}.

@@

Let $t_0$ denote the first event reception time at the tag. Using the pre-measured event period $\tau$ of the RFID reader, the tag can estimate the next event reception time at time $t_1$ as
\begin{eqnarray}
\hat{C}_w(t_1) &=& C_w(t_0)+\tau(1+\hat{f}_w^r(t_0)), \label{eq:core}
\end{eqnarray}
where $\hat{f}_w^r(t_0)$ denotes the \emph{estimated relative frequency} at time $t_0$, which is an estimate for \eqref{eq:relative_frequency}. The intuition behind \eqref{eq:core} is that $1+\hat{f}_w^r(t_0)$ gives the estimated value of $f_r(t_0)/f_w(t_0)$, $\tau f_r(t_0)$ gives the estimated local time passed on the reader side and $\tau f_r(t_0)/f_w(t_0)$ gives the estimated number of clock ticks on the tag side that will pass until the next packet reception. Redefining \eqref{eq:gamma_error}, the \emph{prediction error} at time $t_1$ can be written as
\begin{eqnarray}
\gamma(t_1) & \!\!\!\!\!=\!\!\!\!\!\! & 
C_{w}(t_{1}\!)\!-\!\hat{C}_{w}(t_{1}\!) \!=\! 
C_{w}(t_{1}\!)\!-\!C_{w}(t_{0}\!)\!-\!\tau(1\!\!+\!\!\hat{f}_{w}^{r}(t_0\!)). \label{eq:gamma_t_1}
\end{eqnarray}
Apparently, the objective is to minimize the prediction error so that the event reception time is estimated correctly. As can be observed from \eqref{eq:gamma_t_1}, the estimation error on the relative clock frequency at time $t_0$, i.e. $\hat{f}_w^r(t_0)$, depends on the mean event period $\tau$, and its variance (since they affect $C_w(t_1)-C_w(t_0)$). Inspired by the integral controller update rule in \cite[Section III, (5)]{pi2015}, we propose to update the estimated relative frequency at time $t_1$ as
\begin{eqnarray}
\hat{f}_{w}^{r}(t_1) &=& \hat{f}_{w}^{r}(t_0)+\beta\gamma(t_1), 
\label{eq:update_rule}
\end{eqnarray}
where $\beta$ is the \emph{integral gain}. Assume that the uncertainty of the event period is $\varepsilon\sim\mathcal{N}(0,\sigma^2_{\tau})$ and the instantaneous change on the relative clock frequency is $\frac{df_{w}^{r}(t)}{dt}\sim\mathcal{N}(0,\sigma_{\eta}^{2})$. Under these assumptions, it can be proven that if $0<\beta<\frac{2}{\tau}$ is satisfied the synchronization will hold eventually; i.e. $\hat{f}_{w}^{r}=f_{w}^{r}$. Moreover, the mean and the variance ($\sigma^2_\gamma$) of the steady-state (asymptotic) estimation error of WISP-Sync can be given by 
\begin{eqnarray}
\gamma & \!\!\!\!\!\sim\!\!\!\!\! & \mathcal{N} \!\!\left(\!\!0,\sigma^2_\gamma\!\!=\!\!\frac{\sigma^2_{\eta}\!\left(\!\tau^2\!-\!\beta\tau^3\!+\!\frac{\beta^2\tau^4}{3}\!\right)\!\!+\!\!\beta^2\sigma^2_{\tau}\tau}{(2\beta-\beta^2\tau)}\!\!+\!\!\frac{\sigma^2_{\eta}\tau^3}{3}\!+\!\sigma^2_{\tau} \!\!\right),
\label{eq:sync_performance}
\end{eqnarray}
which is derived in Section A.2 of the Appendix.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/sim-theory.eps}
	\caption{\label{fig:theory-vs-simulations} The plot of the cumulative distribution function (CDF) of $\gamma$ in \eqref{eq:sync_performance} and the CDF of the steady-state $\gamma$ obtained during numerical simulations of \eqref{eq:update_rule} (by implementing (15) and (16) in Appendix). We used $\tau=1$, $\beta= 0.0001$, $\sigma_{\tau}=0.00005$ and our plots depict the theoretical and simulation results with respect to $\sigma_\eta = 0.0002$ and $\sigma_\eta = 0.0008.$ It can be observed that \eqref{eq:sync_performance} is consistent with the simulations and bigger uncertainty leads to bigger synchronization error.}
\end{figure}
In Fig.~\ref{fig:theory-vs-simulations}, we emphasize that our theoretical derivation in \eqref{eq:sync_performance} is consistent with the simulations of \eqref{eq:update_rule}. We refer again the reader to the Appendix for the detailed proofs of the convergence and the asymptotic estimation error. Moreover, in Section~\ref{sec:adaptibility}, we discuss the impact of the parameters $\tau$ and $\beta$ on the synchronization accuracy.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1.4\columnwidth]{figures/alpha-period-vs-variance.eps}
	\caption{\label{fig:simulations} Numerical results showing the relations between event period, integral gain and the steady-state variance. First, we set $\sigma_{\tau}=0.4097$, $\sigma_\eta = 0.0008$, $\beta=1$ and plotted $\sigma^2_{\gamma}$ using different $\tau$ (left hand side). Observe that increasing $\tau$ increases $\sigma^2_{\gamma}$. Then, we set $\sigma_{\tau}=0.4097$,	$\tau=1$ and $\beta \in (0,2)$ due to (19). Using different values for $\sigma_\eta$, we plotted how $\sigma^2_{\gamma}$ changes with respect to the $\beta$ (right hand side). Observe that for each value of $\sigma_\eta$, there is a particular $\beta$ that minimizes $\sigma^2_{\gamma}$.}
\end{figure*}

@@

Since we measured the event period as 226.76\,ms on average (see Fig.~\ref{fig:block_write_delay}), we set $\tau=7086$ 
clock ticks since the local clock of the tag is operating at 32\,kHz and each clock tick occurs every 32 microseconds. First, we explored how $\beta$ affects the performance of the algorithm. By substituting $\tau$ into (19), the convergence condition becomes $0<\beta<0.000284$ in our case. We present the synchronization error with different $\beta$ values in Fig.~\ref{fig:beta}. Observe that the synchronization error tends to get smaller with smaller $\beta$ values but the convergence time increases negligibly. Moreover, due to the low-precision 32\,kHz clock, decreasing $\beta$ has no significant effect after some point. We mention that  in order to make WISP-Sync adapt faster, WISP-Sync can start with a big value of $\beta$ to converge fast and then decrease $\beta$ gradually to decrease the synchronization error, i.e. the adaptive strategy proposed in~\cite{pi2015}. However, from  Fig.~\ref{fig:beta}, we consider a fixed value of $\beta=0.0001$ for the next evaluation steps.

@@

Even though the event-based approach is relatively lightweight and adaptive as compared to the sender-receiver synchronization, it has some limitations. First, the tag is unable to obtain an \emph{explicit reference clock value}. The only variable it tunes is the $\hat{f}_w^r$ that represents the relative frequency of the software clock with respect to the clock of the reader. Therefore, we require additional steps that will allow tags to obtain the actual reference clock value. \add{Moreover, if the tags are covered by different readers, the first step should be synchronizing the readers to the real-time, e.g. by allowing them to access time servers using internet.}
%However, to the best of our knowledge, there is no such command in the EPC Gen 2 standard that allows to send the explicit RFID reader time to the tag.
Second, WISP-Sync requires knowledge about $\tau$, which can be RFID reader dependent and requires a sniffer setup described in Section~\ref{sec:sniffer} to measure. This limitation can be addressed by employing an iterative estimation as an initial step to the WISP-Sync: the tag can start with an initial estimated $\tau$ value, and improve its estimation as it receives successive packets from the reader. After some iterations, the tag can have a rough estimate $\hat{\tau}$ that can be used on behalf of $\tau$. However, estimation errors might also perturb the synchronization accuracy. Last, the steps in Fig.~\ref{fig:event_based} present the communication scenario between one WISP and the reader. To allow other tags sniff this communication and synchronize, a broadcast primitive is crucial---in the next section we show how to overcome this limitation.

@@

%\input{appendix_tmc}
